Linear vertical shear instability in astrophysical disks: a numerical study

### 2014-12-04 linear_iso.tex
starting to add numerical results
added proof on maximum growth rate in limit of weak shear

Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
problem for isothermal perturbations in a vertically isothermal disk, 
\begin{align}\label{iso_ode}
  \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
\end{align} 
subject to appropriate boundary conditions. We remark that Eq
.\ref{iso_ode} is also applicable for    
$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. adiabatic
disks with adiabatic vertical stratification), since in that case  
$W\simeq W$ as well.

@@

>>>>

Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
problem for isothermal perturbations in a vertically isothermal disk, 
\begin{align}\label{iso_ode}
  \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
\end{align} 
subject to appropriate boundary conditions. We remark that Eq
.\ref{iso_ode} is also applicable for    
$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. adiabatic
disks with adiabatic vertical stratification), since in that case  
$Q\simeq W$ as well.

@@

\subsection{Maximum growth rate in the limit of weak shear}   
In the above discussions we have assumed a vertical domain   
$\hat{z}\in[-\infty,\infty]$ which permits analytical expressions for
the growth rates. Here we consider a finite vertical domain
$\hat{z}\in[-\hat{z}_\mathrm{max},\hat{z}_\mathrm{max}]$ and show that
the growth rate is limited by the maximum vertical shear in
the domain. Since we do not persue explicit solutions to the linear
problem, we relax the thin-disk approximation and restore the full
expressions for $\rho$ and $d\Omega^2/dz$.

We begin with the integral relation Eq. \ref{neccessary_cond} with a
finite vertical domain but still assuming boundary terms vanish,
\begin{align}
  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}  
  \notag\\ 
  &=\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}      
  +\ii
  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z},    
\end{align}   
where $f(\hat{z})$ is defined such that
\begin{align}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}
In the thin-disk limit, $f(\hat{z})=\hat{z}$.

We now take the same approach as in the previous section. We first
assume an neutral solution corresponding to $q\equiv0$, for which
$\sigma$ is real and we can take $W$ to be real as well. We then
set $q\to\delta q$ and linearize. This procedure means
\begin{align}
  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}  
  =\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z},\\ 
  &\frac{2\sigma\imag{\delta\sigma}}{\Omega_k^2}\left(1+\hat{k}^2\right)
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}\notag\\
  &=\epsilon\hat{k}\delta q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}. 
\end{align}
Eliminating $\sigma^2$ between the above equations, we obtain
\begin{align}\label{sigma_finite_domain}
  &4
  \left[\frac{\imag{\left(\delta\sigma\right)}}{\Omega_k}\right]^2\left(1+\hat{k}^2\right) 
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}\notag\\
  &=\left(\epsilon\hat{k}\delta q\right)^2\left(\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right)^2. 
\end{align}
On the left hand side of this equation, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  &\left( \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
    |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z}\right)^2\notag\\&\leq
  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho 
  W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
    f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right|\leq \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  |f(\hat{z})| |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z} \notag\\
  \leq
  \mathrm{max}\left(|f|\right)\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
  |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$z\in[-\zmax,\zmax]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}
  \left|\frac{\imag{(\delta\sigma)}}{\Omega_k}\right| \leq
  \frac{\epsilon |\hat{k}\delta
    q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
